TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-07-27 paper/draft.tex
remove old comments
finish sample selection description

% In each case, \thecannon\ produces a predictive model for \apogee\ spectra, but without any input of stellar physics beyond whatever went in to producing the labels in the first place. That is, the model to be applied to the \apogee\ data is learned from the \apogee\ data alone. This obviates physical models that are often inaccurate for M dwarfs, and capitalizes on the excellent data quality delivered by the \apogee\ spectrograph and project. 
%JB: I think this part is redundant

@@

%We note that rotational velocity (vsini) is not a parameter are not included as training labels because that would be hard... also it's probably not needed... I'll let someone else write that. -AWM
%\color{gcolor}{}\color{black}

@@

\begin{figure*}
\begin{center}
	\includegraphics[width=.48\linewidth]{figures/ruwe_cut_before.png}
	\includegraphics[width=.48\linewidth]{figures/ruwe_cut_after.png}
	\caption{}
	\label{fig:ruwe_cut}
\end{center}
\end{figure*}

@@

\item \textbf{Model extraploation cuts:} Because extrapolation with \thecannon\ is unreliable, we select only sources with labels inside the range of the training sample: for the \cannon\ reported temperatures and metallicities we impose $2850<T_{\rm eff}<4150$ and $-0.5<\feh<0.5$, and for \cannon\ reported spectral subtypes we impose $0<SPT<9$.

\item \textbf{Astrometric cut:} \gaia\ re-normalised unit weight error (RUWE) $RUWE<1.2$

\item \textbf{Binary cut:}

@@

The three panels of \ref{fig:safe_selection} show before and after selection of the sources in \gaia\ color-magnitude space, colored by \thecannon\ determined temperatures, metallicities, and spectral types, with their respective training samples overplotted in orange. Each plot shows the expected gradient: temperature increases with decreasing color, spectral subtype increases with increasing color, and metallicity gradient increases perpendicularly to the main sequence branch. Additionally we notice a slight deviation at the upper slice of the main sequence where \cannon\ temperatures, metallicities and spectral types all appear lower than the gradient trends in the rest of the branch, which is most likely the binary sequence. 
% Comparisons between the West-trained spectral types and Mann-trained temperatures and metallicities are shown in Figure \ref{fig:west_vs_mann}: left shows the slight negative correlation between spectral type and $\feh$, and right shows the nonlinear decrease in temperature with visible metallicity gradient. 
We also note that applying our model requires very little computational demand: the time to train and test a model on all 14,827 sources was two minutes on a personal laptop.

@@

% % SUMMARIZE \citealt{Casagrande:2008}, \citealt{Boyajian:2012}, and \citealt{Mann:2015} 
% Shown in the top panels of \ref{fig:teff_comparisons}

@@

% \begin{figure}
% 	\includegraphics[width=\linewidth]{figures/cannon_vs_parsec.png}
% 	\includegraphics[width=\linewidth]{figures/aspcap_vs_parsec1.png}
% 	\caption{Comparison of \cannon\ $\feh$ as compared to metallicity predicted by PARSEC isochrone tracks \citep{Bressan:2012} in \gaia\ color-magnitude space. The two panels show five metallicity tracks (ranging $-0.5<$[M/H]$<+0.5$ in steps of 0.1 dex) at two different ages (3 and 7 Gyr). In the ASPCAP plot only the 8,960 out of 5,875 sources with metallcities $-0.5<$[M/H]$<+0.5$ are shown, while the remaining 1,351 sources with [M/H]$<-0.5$ and [M/H]$>+0.5$.}
% 	\label{fig:isochrone_cmd}
% \end{figure}

% \begin{figure*}
% 	\plotone{figures/aspcap_cannon_label_hist.png}
% 	\caption{Distribution of temperatures and metallicities for 5,875 sources from \thecannon\ compared to the ASPCAP pipeline. Note that \cannon\ temperatures only span $2850<T_{\rm eff}<4150$ and $-0.5<\feh<0.5$\,dex because we have imposed cutoffs for stars that fall outside of the training sample range, however for the same 5,875 stars ASPCAP predictions are skewed higher in temperature and lower in metallicity compared to \cannon\ measurements.}
% 	\label{fig:aspcap_cannon_label_hist} 
% \end{figure*}

@@

% \begin{figure*}[]
% 	\plotone{figures/spt_vs_teff_vs_feh.png}
% 	\caption{West-trained spectral types vs. Mann-trained metallcities and temperatures. For each integer spectral type bin the mean temperatures and metallicities are shown as red error boxes. \label{fig:west_vs_mann}}
% \end{figure*}

%---------------------------

% \begin{figure*}[]
% 	\plotone{figures/mann_radii.png}
% 	\caption{Comparison of radii estimated from M15 relations: on the x-axis radii are derived from  K magnitudes and \cannon\ $\feh$, on the y-axis radii are derived from \cannon\ $\teff$ and $\feh$.}
% 	\label{fig:radii}
% \end{figure*}

>>>>

@@

@@

\begin{figure*}
\begin{center}
	\includegraphics[width=.48\linewidth]{figures/ruwe_cut_before.png}
	\includegraphics[width=.48\linewidth]{figures/ruwe_cut_after.png}
	\caption{The left panel shows the test sample of 8,335 sources after applying Selections $1-3$ described in Section \ref{subsec:test_selection}. The right panel shows the same test sample reduced to 6,221 after applying an astrometric quality cut of ${\rm RUWE}<1.2$ (Selection 4). To further remove sources which were likely binary contamination (Selection 5), we cut out sources above the red line which sparsely lay above the majority of the main-sequence with temperatures and metallicities that deviated from the expected gradient, reducing the final test sample to a size of 5,875 sources. The red line shown is constrained by the coordinates $\left\{ (3.5,12), (1.87,7.5) \right\}$ corresponding to ($BP-RP$, $M_G$).}
	\label{fig:ruwe_cut}
\end{center}
\end{figure*}

@@

\item \textbf{Model extraploation cuts:} Because \thecannon\ does not perform well under extrapolated regions of parameter space, we select only sources inside the range of our training sample with ASPCAP parameters $2800<T_{\rm eff}<4100$ and $-0.5<[M/H]<0.5$, and with \cannon\ parameters of $2850<T_{\rm eff}<4150$, $-0.5<\feh<0.5$, and $0<SPT<9$.

\item \textbf{Astrometric cut:} Using the \gaia\ re-normalised unit weight error (RUWE), a metric of evaluating the fit of the astrometric solution described in the additional release notes, we apply a cut of ${\rm RUWE}<1.2$ to select out sources with high astrometric error or noise, such as binaries (see Figure \ref{fig:ruwe_cut}).

\item \textbf{Binary cut:} To remove further contamination from binary sources by a conservative cut, we applied an additional color-magnitude selection on sources above the main-sequence with \cannon\ temperatures/metallicities that deviated from the gradient, which we visibly selected for in Figure \ref{fig:ruwe_cut}.

@@

The three panels of \ref{fig:safe_selection} show before and after selection of the sources in \gaia\ color-magnitude space, colored by \thecannon\ determined temperatures, metallicities, and spectral types, with their respective training samples overplotted in orange. Each plot shows the expected gradient: temperature increases with decreasing color, spectral subtype increases with increasing color, and metallicity gradient increases perpendicularly to the main sequence branch. 
We also note that applying our model requires very little computational demand: the time to train and test a model on all 14,827 sources was two minutes on a 2.7\,GHz Intel core i7 laptop.

@@

@@

@@
